---
title: "autopolate Notebook"
output: html_notebook
---

How to use autopolate package : 

First we intall the package from github using devtools::install_github :

```{r}
# First we make sure that devtools is installed on your R envirnoment
if (!require(devtools)) {
  install.packages('devtools')
}  
# Then we can install my script by running the following command
devtools::install_github('Moustapha-A/autopolate')
```

Then we load the library as any other library

```{r}
library(autopolate)
```

We read some data, CNBC is an example data included in this package you must load your own data frame here

```{r}
data(CNBC)
```

The first rows of the data

```{r}
print(head(CNBC,10))
```

The current data rate is 1 min we want to transform it into 30 seconds and for only the tempreture attribute
This might take time as it is interpolationg a set of 5555 values (I will work on performance if I had time)

```{r}
interpolatedTemp= autopolate(dataframe = CNBC,
                  timeCol = 'Time',
                  timeFrmt = '%d/%m/%Y %H:%M:%S',
                  valueCol = 'Temp',
                  breaksGen = 'gap-aware',
                  rate = 60,
                  targetRate = 30)
```

The first rows of the interpolated tempreture we can see the rate in the Time attribute

```{r}
print(head(interpolatedTemp,10))
```






