---
title: "Auto Interpolation Library For R"
output: html_notebook
---

```{r}
# First we make sure that devtools is installed on your R envirnoment
if (!require(devtools)) {
  install.packages('devtools')
}  
# Then we can install my script by running the following command
devtools::install_github('Moustapha-A/autopolate')
```
First we intall the package from github using devtools::install_github


```{r}
library(autopolate)
```
Then we load the library as any other library

***

### *Treating Neat Data*

```{r}
data("EZ68")
EZ68 = na.omit(EZ68)
```
Now we will load some data that belongs to one of polluscope's campaigns. Using the Ecomzen 68 sensor.

```{r}
head(EZ68,10)
```
The first 10 records of EZ68

```{r}
plot(as.POSIXct(EZ68$`Date (UTC)`,format="%Y-%m-%d %H:%M:%S"),EZ68$NO2, main = "Ecomzen NO2 TS", xlab = "Time", ylab = "NO2")
```
The plot of NO2 against Time

We can see that the aquisition rate of Ecomzen is not stable sometimes it is ~1 min other is ~2 min. What we want to do is to have the values at a stable rate. Let us say that we need to have the data as a rate of 30 seconds. We will do that using the *autopolate* function I developed:

```{r}
NO2= autopolate(dataframe = EZ68,
                  timeCol = "Date (UTC)",
                  timeFrmt = "%Y-%m-%d %H:%M:%S",
                  valueCol = "NO2",
                  targetRate = 30)
```


